---
title: "Covid Project Group 13"
output:
  html_document:
    number_sections: yes
    toc: yes  
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}

library(dplyr)
library(ggplot2)
library(tidyverse)
library(janitor)
library(gganimate)
library(gifski)
library(png)
library(ggrepel)
library(scales)
library(RCurl)

combined_data = read.csv(file = "combinedData.csv")

# obtain up to date data for daily cases automatically
timeline <- read.csv(text = getURL(
  "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_US.csv"
  ))
```

### Group 13's Goals

Our group chose to analyze two different trends in the data from JHU's Covid Dataset, both trying to evaluate correlation, but one looking at data over time, and the other as a summary to date.

1. For our first hypothesis we wanted to test if major holidays or turning points contributed to a rise in the number of cases of coronavirus in the US.

2. For our second hypothesis we wanted to analyze how a region's political affiliation impacted an individual's likelihood to have contracted or been exposed to coronavirus.

## Important Days and Case Increases

Overview ...

Do certain Holiday celebrations line up with increased rates of spread of Covid-19?

First we will create a graph of the new cases each day in the US.

```{r}

# first get a list of just the total cases each day
cases <- timeline %>% select(seq(12, length(timeline)))
totals <- cases %>% summarize_at(vars(41:length(cases)), sum)
total_list <- data.matrix(totals)

# convert from total cases to new cases
new_cases <- c()
for (i in seq(2, length(total_list))) {
  new_cases <- c(new_cases, (total_list[i] - total_list[i-1]))
}
plot(new_cases, type="l", main="Daily New Covid-19 Cases in the US", xlab = "Days since March 1, 2020", ylab = "New Cases")

```

Now Let's take a look at the rate of change of this graph to see where the spread sped up and slowed down.

To smooth the plot out, we will measure the rate of change as the difference between the total cases for the seven days after a given date and the total cases for the seven days before this date.

Black points are dates on which the weekly total cases started increasing

Red lines are labeled with specific holidays

```{r}
# we can also plot this graph's derivative to see where the number of new cases began increasing
# plotting differences in day to day new cases isn't useful since the rate of change oscillates daily
# let's plot differences between 1 week intervals to see more long term trends
rate_of_change <- c()
for (i in seq(8, length(new_cases) - 7)) {
  rate_of_change <- c(rate_of_change, sum(new_cases[(i):(i+7)]) - sum(new_cases[(i-7):(i)]))
}

# since i started at 8, the new graph starts at March 8 instead of March 1.
start_date <- "2020-03-08"

plot(rate_of_change, type="l", main = "Change in weekly total of US Covid-19 Cases", xlab = "Days Since March 8, 2020", ylab = "Change in Weekly Total Cases")
abline(h=0)

# create A line and label for each holiday
# Halloween, election day, and veterans day all occur in a 2-week time period, so we group them together starting from Halloween
# Note that case numbers following Thanksgiving may not be fully updated by the time of the presentation
holidays <- c("Easter", "Memorial Day", "Independence Day", "Labor Day",  "Halloween", "Thanksgiving")
dates <- c("2020-04-12", "2020-05-25","2020-07-04", "2020-09-07",  "2020-10-31", "2020-11-26")
height <- 200000
for (i in seq(length(dates))) {
  abline(v=as.Date(dates[i])-as.Date(start_date), col="red")
  text(x = as.Date(dates[i])-as.Date(start_date), y = height, holidays[i], adj = c(1, NA))
  # Alternate the heights of the labels
  if (height == 200000) {
    height <- 150000
  } else {
    height <- 200000
  }
}

# find the dates where the rate of change became positive
# need to start off with a date to get started, will remove after loop
zeroes <- as.Date(start_date)
dates <- seq(from=as.Date(start_date), by="day", length.out=length(totals))
for (i in seq(2, length(rate_of_change))) {
  if (rate_of_change[i] > 0 && rate_of_change[i-1] < 0) {
    zeroes <- c(zeroes, dates[i])
  }
}
# remove the starting date
zeroes <- zeroes[2:length(zeroes)]

# plot points for the zeroes
zeroes <- zeroes - as.Date(start_date)
for (z in zeroes) {
  points(z, y=0, lwd = 5)
  print(as.numeric(z))
}

```

Looking at the change in weekly cases, there are a few holidays such as Memorial Day
and Labor Day which seem to be followed by an increased rate of spread. However,
other days such as Easter and Independence Day saw a minimal change or even a decrease
in the rate of spread. Also interesting to note is the dip just before Thanksgiving,
which may reflect a change in testing than a change in actual cases. The Holidays do
seem to line up with interesting points on the graph, but they are not necessarily
the cause of these changes.

## Political Affilitation and Likelihood of Contracting Coronavirus

Overview ...

- Scatter Plot
- Linear Regression
- Clustering

```{r}
#normalize cases by dividing by the population of each county
combined_data <- transform(combined_data, cases_per_pop = as.integer(Cases) / as.integer(Population))
cases_per_pop <- combined_data$cases_per_pop
percent_gop <- combined_data$per_gop
percent_dem <- combined_data$per_dem


#all data should be between 0 and 1, make sure it is
#if there are random outliers and data is correct!! add ylim=c(0,1) to both graphs
gop_vs_cases <- lm(cases_per_pop ~ percent_gop)
summary(gop_vs_cases) #use this to check p value!!
plot(percent_gop, cases_per_pop, col = "red", xlab = "Percentage of Population Voting Republican", ylab = "Cases as Percent of Population")
abline(gop_vs_cases, col ="blue")


dem_vs_cases <- lm(cases_per_pop ~ percent_dem)
summary(dem_vs_cases) #use this to check p value!!
plot(percent_dem, cases_per_pop, col = "blue", xlab = "Percentage of Population Voting Democrat", ylab = "Cases as Percent of Population")
abline(dem_vs_cases, col = "red")

# correlation values between percent gop, dem and cases/pop
final = data.frame("cases.pop" = cases_per_pop, "pecent.dem" = percent_dem, "percent.gop" = percent_gop)

cor(final, use = "complete.obs")

```

Based on the R-squared values alone, it would look like there is significant
correlation between these variables. However, the slope of this linear relationship
is very shallow, so this is not a strong correlation. Additionally, we can see
a large number of outliers from the linear relationship for counties with
a high proportion of Republican voters, which may have contributed to this slope.
It is possible there are other variables affecting this relationship rather than
a direct correlation between political affiliation and proportion of covid cases.
