---
title: "Covid Project Group 13"
output:
  html_document:
    number_sections: yes
    toc: yes  
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
# we can use this section here to load any packages that will be needed
knitr::opts_chunk$set(echo = TRUE)
covid_and_population_summary = read.csv(file = "casesPopbyCounty.csv")
election_data = read.csv(file = "clean2020electiondata.csv")
combined_data = read.csv(file = "combinedData.csv")
```

### Group 13's Goals

Our group chose to analyze two different trends in the data from JHU's Covid Dataset, both trying to evaluate correlation, but one looking at data over time, and the other as a summary to date.

1. For our first hypothesis we wanted to test if major holidays or turning points contributed to a rise in the number of cases of coronavirus in the US.

2. For our second hypothesis we wanted to analyze how a region's political affiliation impacted an individual's likelihood to have contracted or been exposed to coronavirus.

## Important Days and Case Increases

Overview ...

Following Key dates, is there a spike in the following 2-5 days after these events?

- Timeline
- Key dates
- Correlation

```{r}
#code, graphs, etc
```

Key takeaways

## Political Affilitation and Likelihood of Contracting Coronavirus

Overview ...

- Scatter Plot
- Linear Regression
- Clustering

```{r}
# Remove Outliers (remove if outliers are wanted)
cases <- combined_data$Cases
cases_Q1 <- quantile(cases, 0.25)
cases_Q3 <- quantile(cases, 0.75)
cases_IQR <- IQR(cases)
# adjust 1.5 to 3 to remove only extreme outliers
combined_data_no_out <- subset(combined_data, cases> (cases_Q1 - 1.5*cases_IQR) & cases< (cases_Q3 + 1.5*cases_IQR))

# Grab data from sets (change combined_data_no_out to combined_data to use set with outliers)
cases <- combined_data_no_out$Cases
percent_gop <- combined_data_no_out$per_gop
percent_dem <- combined_data_no_out$per_dem

# Plot the scatterplots (seperately, they appear too squished if together)
plot(percent_gop, cases, col = "red", xlab = "Percentage of Population Voting Republican", ylab = "Number of Cases")
plot(percent_dem, cases, col = "blue", xlab = "Percentage of Population Voting Democrat", ylab = "Number of Cases")
```

Key takeaways

## References

Source out all our data
