---
title: "Covid Project Group 13"
output:
  html_document:
    number_sections: yes
    toc: yes  
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(tidyverse)
library(janitor)
library(gganimate)
library(gifski)
library(png)
library(ggrepel)
library(scales)

```

### Group 13's Goals

Our group chose to analyze two different trends in the data from JHU's Covid Dataset, both trying to evaluate correlation, but one looking at data over time, and the other as a summary to date.

1. For our first hypothesis we wanted to test if major holidays or turning points contributed to a rise in the number of cases of coronavirus in the US.

2. For our second hypothesis we wanted to analyze how a region's political affiliation impacted an individual's likelihood to have contracted or been exposed to coronavirus.

## Important Days and Case Increases

Overview ...

Following Key dates, is there a spike in the following 2-5 days after these events?

- Timeline
- Key dates
- Correlation

```{r}
#code, graphs, etc
library(dplyr)
library(ggplot2)
library(tidyverse)
library(janitor)
library(gganimate)
library(gifski)
library(png)
library(ggrepel)
library(scales)
# we can use this section here to load any packages that will be needed
knitr::opts_chunk$set(echo = TRUE)
covid_and_population_summary = read.csv(file = "casesduringholidaysCorrected.csv")
head(covid_and_population_summary)

covid_full_1 <- covid_and_population_summary %>% select(c(1:6)) 
covid_map_state_1 <- covid_full_1 %>% group_by(Holidays) %>% 
  summarize_at(vars(1:3), sum) %>% adorn_totals('row')
covid_map_state_1$Province_State <- tolower(covid_map_state_1$Province_State)


```

Key takeaways

## Political Affilitation and Likelihood of Contracting Coronavirus

Overview ...

- Scatter Plot
- Linear Regression
- Clustering

```{}

```

Key takeaways

## References

Source out all our data
